On The Subspace Learning for Network Attack Detection

### 2019-05-27 docs/thesis/chapters/3_mobile/mobile.tex
first deliverable

Cloud computing is a new rapidly evolving paradigm in the world of distributed networking and computation. The basic features of the cloud environment is providing the elastic, on-demand and secure services for the end-users. While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}. Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized leading to obscure and incomplete security paradigms and solutions.

A common practice to provide a cloud solution with stable security is to use a specific type of cloud service, such as Cloud Access Security Broker (CASB) or Cloud Access Control (CAC). These services are specifically designed to bring security at a single access point and provide the coordination of the most important security measures. Gartner \cite{skyhigh2015} estimates that such systems will be used by 85 \%  of companies by 2020. The reason for this is that the organization of security measures at a single control point allows to control and to monitor the level of cloud protection much more effectively. The basic features of the CASB are: discovery of cloud services, encryption along with tokenization for better search properties, access control, Data Loss Prevention (DLP) services, authentication, and auditing and alerting services \cite{lawson2015}. The protection of the confidential data, according to the standards of CASB deployment should be provided elsewhere, i.e. in transit and at the end-user, i.e. end-client, side \cite{campbell2015}.

@@

The protection scheme used on a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}. Therefore, encrypting files and generating keys on a mobile device is not considered a good solution. On the other hand, the protection schemes with good computational qualities lack the security analysis in many cases \cite{khan2014bss}. The common practice is the shadow user activities monitoring \cite{yovel2014}. However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode.

Suppose that a SME uses CASB to protect data at rest, i.e. while stored on the server-side, in transit, communicating with server, and, in use, while the mobile client is connected to the network. In this case, when the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile client has to secure itself with its own limited resources. Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection. For example, offline mode does not allow performing the extensive computation and encryption on the mobile device.
Observing the above described landscape, this work outlines the concept of the offline mobile cloud security.

This work proposes an architecture and approach for user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack. The key expiration period is safely incorporated into the proposed system solution in order to enhance security. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.

The work is structured as follows. Section \ref{sec:3_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions for offline protection in the BYOD world. Section \ref{sec:3_the_client_security_architecture} outlines the mobile security architecture of the proposed solution. Section \ref{sec:3_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section \ref{sec:3_mos} explains the algorithms for MOS. Section \ref{sec:3_results} discusses the common threat scenarios, the data modeling, the performance analysis and discussion of the practical implementation results. Section \ref{sec:3_conclusion} concludes the chapter.

@@

In many cases the industrial providers of the mobile security API avoid the problem by completely forbidding the offline access to the protected app, i.e. SAP Mocona, which operates as a secure client wrapping layer \cite{Lelyveld2013}. Due to various constraints such as traffic load, traveling and ease of access, the SME business procedure may require such access.

@@

Figure \ref{fig:3_02} depicts the client-side protection mechanisms. The client should support 4 subsystems: i) encryption subsystem (\textbf{Online/Offline Cryptography}) that contains the procedures of encryption and decryption; ii) storage subsystem (\textbf{Protected Storage}) that provides the downloaded shares and key storage protection; iii) (\textbf{threat intelligence}) unit that provides the constant monitoring, and iv) the communication subsystem (\textbf{Communication with Server}). In short, all security procedures are connected to 4 groups of operations: file request and receiving; encryption and decryption; file and key storing; monitoring and analysis.

@@

The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and MOS-inspired statistics analyzer. Functions such as alerting and deleting the expired key belong to this module as well. These functions are described in Subsection \ref{sec:3_mos}.

@@

\begin{enumerate}
	\item \textbf{Protected Storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files.
	\item \textbf{Threat Intelligence Manager (TIM)}: most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode. Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios. Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior, according to the scenarios in Section \ref{sec:3_results}. Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones. The TIM is an internal module of the mobile client that implements offline anomaly detection through signal processing techniques.
	\item \textbf{Key Management Center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
\end{enumerate}

@@

In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, perform feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode. Figure \ref{fig:3_06} depicts the TIM for offline behavioral analysis.

@@

The LAC performs the behavioral analysis through eigenvalue analysis and MOS schemes, which identify anomalies on sparse, subtle or abrupt number of user operations. The malicious behavior detection is detail described in \ref{sec:3_mos}.

\section{MOS for Threat Intelligence}
\label{sec:3_mos}
In the context of anomaly-based schemes for attack detection, the proposed behavioral analysis approach applies signal processing techniques, such as Principal Component Analysis and Model Order Selection schemes \cite{tenorio2013greatest}, for automatic detection of attacks or malicious behaviors.

Model Order Selection is an effective signal processing technique for several applications, allowing separating the only noise components from the principal components applying a rank reduction of the data.

Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. MOS for threat intelligence requires that the target features, which can be user operations, should be modeled as a matrix composed by the number of occurrences by location and time, and grouped into $Q$ time frames. Therefore, the framework considers the time variations of the matrix $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the occurrence of malicious behaviors. For example, one element of $\mathbf{X}^{(q)}$ can represent the number of file readings on folder $m$ during the minute $n$, from file operations logged by the mobile client.

The MOS schemes can rely on sample covariance of zero mean variables (called as \textbf{zero mean covariance} for the sake of simplicity) and sample covariance of zero mean and unitary standard deviation (called as \textbf{zero mean and standardized covariance} for the sake of simplicity) variables, where the former is useful to identify abnormalities caused by large amounts of operations during a period, while the latter is applied to identify anomalies on sparse or subtle number of file operations.

@@

\begin{equation}\label{eq:3_eq03}
	\mathbf{y}_{m}^{(q)} = \mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}.
\end{equation}

@@

\begin{equation}\label{eq:3_eq04}
	\mathbf{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\mathbf{Y}^{(q)}\mathbf{Y}^{(q)^{\rm T}},
\end{equation}
where $\mathbf{\hat{R}}_{yy}^{(q)}$ means the estimation of the sample covariance matrix from the measured zero mean matrix $\mathbf{Y}^{(q)}$ over $N$ minutes of the time frame $q$.

For MOS based on sample covariance of zero mean and unitary standard deviation, in order to identify anomalies with sparse or subtle behavior, it is required, for each variable, to make the standard deviation unitary as follows

\begin{equation}\label{eq:3_eq05}
	\mathbf{z}_{m}^{(q)} = \frac{\mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}}{\mathbf{\sigma}_{m}^{(q)}}.
\end{equation}

@@

\begin{equation}\label{eq:3_eq06}
	\mathbf{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\mathbf{Z}^{(q)}\mathbf{Z}^{(q)^{\rm T}}.
\end{equation}

Once the $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ have been obtained for MOS in order to anomaly detection, for the sake of simplicity, we refer to $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ as a matrix $\mathbf{C}$. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\mathbf{C}^{(q)} = \mathbf{V}^{(q)}\mathbf{\lambda}^{(q)}\mathbf{V}^{(q)^{\rm T}}$, in order to obtain the vector of eigenvalues $\textbf{e}$, as following:

\begin{equation}\label{eq:3_eq07}
	\mathbf{e}^{(q)} = \rm diag(\mathbf{\lambda}^{(q)}),
\end{equation}

@@

The process of obtaining the $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$ and the matrix $\mathbf{C}^{(q)}$, finding the largest eigenvalue for each $q$-th time frame, should be repeated until $q = Q$, in order to obtain the largest eigenvalue of all time frames, as presented by

\begin{equation}\label{eq:3_eq08}
	\mathbf{E} =
	\begin{bmatrix}
		\lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
		\lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
		\lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
		\vdots & \vdots & \ddots & \vdots  \\
		\lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
	\end{bmatrix}.
\end{equation}

@@

\begin{equation}\label{eq:3_eq09}
	\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} \cdots \lambda_1^{(Q)}]
\end{equation}

@@

\begin{equation}\label{eq:3_eq10}
	\hat{d} = \textrm{MOS}(\boldsymbol{e}_{\rm max})
\end{equation}

@@

MOS schemes based on zero mean covariance analysis are effective to reveal abrupt changing of behaviors over time \cite{tenorio2013greatest}, making possible to identify intense malicious behaviors on offline mode of mobile clients, such in case of ransomware attack or bulk access to sensitive data.

@@

Sparse or subtle file operations, with low number of operations distributed over different files or directories, during short period of time can indicate anomalies in contrast to the required time for legitimate directory navigation. MOS and zero mean and standardized covariance analysis can be suitable if applied to evaluate the time and location of operations, in order to identify unreachable navigation, if compared to legitimate navigation

The MOS based on zero mean covariance analysis indicates abnormalities caused by large amounts of operations during a period. Subsequently, the eigenvalue analysis highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.

@@

\begin{enumerate}[label=(\alph*)]
	\item The hacker has access to the mobile client and is able to perform operations;
	\item The session time is valid;
	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
	\item The mobile client incrementally append each operation attempt time into the logging;
	\item The MOS module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
	\item The MOS module evaluates the logging of legitimate operations, applying zero mean covariance analysis to identify abnormalities caused by massive operations during the session time.
\end{enumerate}

@@

In the offline mode, the session time is used to restrict the operations during a specified period, although it is possible to manipulate the current time in mobile device, to emulate a period in which the session was valid. The log analysis by MOS can deal with this kind of threat, through the incremental logging of the time when each operation was performed, followed by the behavioral evaluation of operations over time.

@@

Applying MOS to the analysis of the time between user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. The MOS based on zero mean and standardized covariance analysis identifies anomalies on sparse or subtle variation in the number of file operations, since the eigenvalue analysis highlights the unexpected number of sparse (such as file operations on diverse folders) or subtle operations. Consequently, the result of the eigenvalue analysis is applied to MOS schemes, in order to identify the occurrence of malicious behaviors during the valid session.

The MOS based on zero mean covariance analysis indicates abnormalities caused by large amounts of operations during a period. The eigenvalue analysis based on the zero mean covariance matrix highlights massive or concentrated operations over time or location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.

@@

\begin{enumerate}[label=(\alph*)]
	\item The hacker steals the operating system;
	\item The hacker modifies the time of the operating system to a period when the session was valid;
	\item The hacker has access to the mobile client and is able to perform operations;
	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
	\item The mobile client incrementally append each operation attempt time into the logging;
	\item The mobile client verifies if one logged time is older than the last operation time. If it is true, the MOS module classifies the evaluated operation as malicious;
	\item The MOS module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations;
	\item The MOS module evaluates the logging of legitimate operations, applying zero mean covariance analysis to identify abnormalities caused by massive operations during the session time;
\end{enumerate}

@@

MOS schemes are used in order to identify anomalous behavior that can indicate an attack and be used to prevent or avoid attack propagation. Therefore, it is necessary to analyze the data that can be collected from user operations on mobile client, to identify features that can be modeled and submitted to MOS schemes, according to described in Section \ref{sec:3_mos}.

The selected features shall be modeled as matrices which represents a signal superposition containing noise, legitimate and malicious behavior \cite{tenorio2013greatest}, grouped into time frames $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, 2, 3, \ldots, Q$ , where $M$ defines the decomposition of a selected feature, $N$ defines the time decomposition and represents the number of occurrences of the feature $m$ during the time $n$.

@@

An important security issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile clients. Once a mobile client releases the connection with the corporate cloud, no security measure implemented in the cloud infrastructure assures the protection of sensitive data stored in the mobile device. Aware of this problem and its importance, this work presented a proposal to address the offline mobile security problem combining a different cryptographic methods. Moreover, this proposed approach also prevents malicious user behavior by applying a MOS-based analytic method.

@@

Future works in the area can further explore enhancements in the analytics methods as well as to extended the approach to be evaluate multidimensional data through tensor-based analysis.

>>>>

The protection scheme used in a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}. Therefore, encrypting files and generating keys on a mobile device is not considered a good solution. On the other hand, the protection schemes with good computational qualities lack the security analysis in many cases \cite{khan2014bss}. The common practice is the shadow user activities monitoring \cite{yovel2014}. However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode. When the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools can not help and the mobile client has to secure itself with its own limited resources. Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection.

@@

This work proposes an architecture and approach for user behavior analysis based on Eigensim \cite{vieira2017model}, in order to detect possible threats in offline corporate mobile applications. The key expiration period is safely incorporated into the proposed system solution in order to enhance security and the behavioral analysis can indicate malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.

The work is structured as follows. Section \ref{sec:3_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions for offline protection in the BYOD world. Section \ref{sec:3_the_client_security_architecture} outlines the mobile security architecture of the proposed solution. Section \ref{sec:3_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section \ref{sec:3_eigensim} explains the use of Eigensim. Section \ref{sec:3_results} discusses the common threat scenarios, the data modeling, the performance analysis and discussion of the practical implementation results. Section \ref{sec:3_conclusion} concludes the chapter.

@@

@@

Figure \ref{fig:3_02} depicts the client-side protection mechanisms. The client should support 4 subsystems: 
\begin{enumerate}
	\item Encryption subsystem that provides the procedures of encryption and decryption;
	\item Protected storage subsystem that provides the downloaded shares and key storage protection;
	\item The Threat intelligence unit that provides the constant monitoring;
	\item The communication subsystem that enables with the server.
\end{enumerate}

In summary, all security procedures are connected to 4 groups of operations: file request and receiving; encryption and decryption; file and key storing; monitoring and analysis.

@@

The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and Eigensim-inspired statistics analyzer. Functions such as alerting and deleting the expired key belong to this module as well. These functions are described in Subsection \ref{sec:3_eigensim}.

@@

\begin{enumerate}
	\item \textbf{Protected Storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files.
	\item \textbf{Threat Intelligence Manager (TIM)}: most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode. Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal,vieira2017model} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios. Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior, according to the scenarios in Section \ref{sec:3_results}. Moreover, the Eigensim, which is an approach based on subspace learning and on effective signal processing technique to separate noise components from the principal components, named Model Order Selection (MOS), can be applied into anomaly and attack detection \cite{tenorio2013greatest, vieira2017model}, to identify and separate malicious behaviors from the legitimate ones. The TIM is an internal module of the mobile client that implements offline anomaly detection through signal processing techniques.
	\item \textbf{Key Management Center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
\end{enumerate}

@@

In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, perform feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode. Figure \ref{fig:3_06} depicts the TIM workflow for offline behavioral analysis.

@@

The LAC performs the behavioral analysis through eigenvalue analysis and MOS schemes proposed by Eigensim, which identify anomalies on sparse, subtle or abrupt number of user operations. The malicious behavior detection is detail described in \ref{sec:3_eigensim}.

\section{Eigensim for Threat Intelligence}
\label{sec:3_eigensim}
In the context of anomaly-based schemes for attack detection, the proposed behavioral analysis approach applies signal processing techniques, such as subspace learning by eigenvalue decomposition and Model Order Selection schemes \cite{tenorio2013greatest}, for automatic detection of attacks or malicious behaviors. Model Order Selection is an effective signal processing technique for several applications, allowing separating the only noise components from the principal components applying a rank reduction of the data.

Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. MOS for threat intelligence requires that the target features, such as user operations, should be modeled as a matrix composed by the number of occurrences grouped by location and time, and splited into $Q$ time frames. Therefore, the framework considers the time variations of the matrix $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the occurrence of malicious behaviors. For example, one element of $\mathbf{X}^{(q)}$ can represent the number of file readings on folder $m$ during the minute $n$, from file operations logged by the mobile client.

The Eigensim can rely on sample covariance of zero mean variables (called as \textbf{zero mean covariance} for the sake of simplicity) and sample covariance of zero mean and unitary standard deviation (called as \textbf{zero mean and standardized covariance} for the sake of simplicity) variables, where the former is useful to identify abnormalities caused by large amounts of operations during a period, while the latter is applied to identify anomalies on sparse or subtle number of file operations.

@@

\begin{equation}\label{eq:3.03}
	\mathbf{y}_{m}^{(q)} = \mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}.
\end{equation}

@@

\begin{equation}\label{eq:3.04}
	\mathbf{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\mathbf{Y}^{(q)}\mathbf{Y}^{(q)^{\rm T}},
\end{equation}
where $\mathbf{\hat{R}}_{yy}^{(q)}$ means the estimation of the sample covariance matrix from the measured zero mean matrix $\mathbf{Y}^{(q)}$ over $N$ minutes of the time frame $q$.

For Eigensim based on sample covariance of zero mean and unitary standard deviation, in order to identify anomalies with sparse or subtle behavior, it is required, for each variable, to make the standard deviation unitary as follows

\begin{equation}\label{eq:3.05}
	\mathbf{z}_{m}^{(q)} = \frac{\mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}}{\mathbf{\sigma}_{m}^{(q)}}.
\end{equation}

@@

\begin{equation}\label{eq:3.06}
	\mathbf{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\mathbf{Z}^{(q)}\mathbf{Z}^{(q)^{\rm T}}.
\end{equation}

Once the $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ have been obtained for anomaly detection based on Eigensim, for the sake of simplicity, we refer to $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ as a matrix $\mathbf{\hat{R}}$. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\mathbf{\hat{R}}^{(q)} = \mathbf{V}^{(q)}\mathbf{\lambda}^{(q)}\mathbf{V}^{(q)^{\rm T}}$, in order to obtain the vector of eigenvalues $\textbf{e}$, as following:

\begin{equation}\label{eq:3.07}
	\mathbf{e}^{(q)} = \rm diag(\mathbf{\lambda}^{(q)}),
\end{equation}

@@

The process of obtaining the $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$ and the matrix $\mathbf{\hat{R}}^{(q)}$, finding the largest eigenvalue for each $q$-th time frame, should be repeated until $q = Q$, in order to obtain the largest eigenvalue of all time frames, as presented by

\begin{equation}\label{eq:3.08}
	\mathbf{E} =
	\begin{bmatrix}
		\lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
		\lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
		\lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
		\vdots & \vdots & \ddots & \vdots  \\
		\lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
	\end{bmatrix}.
\end{equation}

@@

\begin{equation}\label{eq:3.09}
	\boldsymbol{e}_{\rm max} = [ \lambda_1^{(1)}, \lambda_1^{(2)} \cdots \lambda_1^{(Q)}]
\end{equation}

@@

\begin{equation}\label{eq:3.10}
	\hat{d} = \textrm{MOS}(\boldsymbol{e}_{\rm max})
\end{equation}

@@

The Eigensim and MOS schemes based on zero mean covariance analysis are effective to reveal abrupt changing of behaviors over time \cite{tenorio2013greatest}, making possible to identify intense malicious behaviors on offline mode of mobile clients, such in case of ransomware attack or bulk access to sensitive data.

@@

Sparse or subtle file operations, with low number of operations distributed over different files or directories, during short period of time can indicate anomalies in contrast to the required time for legitimate directory navigation. Eigensim and zero mean and standardized covariance analysis can be suitable if applied to evaluate the time and location of operations, in order to identify unreachable navigation, if compared to legitimate navigation

The Eigensim based on zero mean covariance analysis indicates abnormalities caused by large amounts of operations during a period. Subsequently, the eigenvalue analysis highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.

@@

\begin{enumerate}[label=(\alph*)]
	\item The hacker has access to the mobile client and is able to perform operations;
	\item The session time is valid;
	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
	\item The mobile client incrementally append each operation attempt time into the logging;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean covariance analysis to identify abnormalities caused by massive operations during the session time.
\end{enumerate}

@@

In the offline mode, the session time is used to restrict the operations during a specified period, although it is possible to manipulate the current time in mobile device, to emulate a period in which the session was valid. The log analysis by Eigensim can deal with this kind of threat, through the incremental logging of the time when each operation was performed, followed by the behavioral evaluation of operations over time.

@@

Applying Eigensim to the analysis of the time between user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. The MOS based on zero mean and standardized covariance analysis identifies anomalies on sparse or subtle variation in the number of file operations, since the eigenvalue analysis highlights the unexpected number of sparse (such as file operations on diverse folders) or subtle operations. Consequently, the result of the eigenvalue analysis is applied to MOS schemes, in order to identify the occurrence of malicious behaviors during the valid session.

The Eigensim based on zero mean covariance analysis indicates abnormalities caused by large amounts of operations during a period. The eigenvalue analysis based on the zero mean covariance matrix highlights massive or concentrated operations over time or location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.

@@

\begin{enumerate}[label=(\alph*)]
	\item The hacker steals the operating system;
	\item The hacker modifies the time of the operating system to a period when the session was valid;
	\item The hacker has access to the mobile client and is able to perform operations;
	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
	\item The mobile client incrementally append each operation attempt time into the logging;
	\item The mobile client verifies if one logged time is older than the last operation time. If it is true, the Eigensim module classifies the evaluated operation as malicious;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean covariance analysis to identify abnormalities caused by massive operations during the session time;
\end{enumerate}

@@

The Eigensim and MOS schemes are used in order to identify anomalous behavior that can indicate an attack and be used to prevent or avoid attack propagation. Therefore, it is necessary to analyze the data that can be collected from user operations on mobile client, to identify features that can be modeled and submitted to Eigensim, according to described in Section \ref{sec:3_eigensim}.

The selected features shall be modeled as matrices which represents a signal superposition of signal, artifact and noise, which refer to normal, controlled and anomalous behaviors \cite{tenorio2013greatest,vieira2017model}. Thus, the data is grouped into time frames $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, 2, 3, \ldots, Q$ , where $M$ defines the decomposition of a selected feature, $N$ defines the time decomposition and represents the number of occurrences of the feature $m$ during the time $n$.

@@

An important security issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile clients. Once a mobile client releases the connection with the corporate cloud, no security measure implemented in the cloud infrastructure assures the protection of sensitive data stored in the mobile device. Aware of this problem and its importance, this work presented a proposal to address the offline mobile security problem combining cryptographic methods and a Eigensim-based behavioral anomaly detection.

@@
